Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-06 Manuscript/manuscript.tex
Written/edited part dealing with P-N vector and COO- vector in the section with calcium. Next: Continue editing the part with the interacting moieties.

@@

%%% Present the density profiles (different concentrations for both POPC and mixed)
The increase in the amount of bound calcium cations from pure POPC \cite{melcr18}
to the mixed negatively charged bilayer containing POPS
is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
summarized in Table~\ref{tab:binding}. 
Distributions of \ce{Ca^{2+}}, \ce{Na^+} counterions and also \ce{Cl^-} 
are plotted in Fig.~\ref{fig:cacl-dens_PCPS} for the negatively charged bilayer. 
In contrast to \ce{KCl} or added concentrations of \ce{NaCl}, 
\ce{Na^+} counterions are substituted with \ce{Ca^{2+}} even at low concentrations of \ce{CaCl2}. 
The increasing concentration of \ce{CaCl2} and, hence, a higher amount of bound \ce{Ca^{2+}}
also attracts \ce{Cl^-} anions to the bilayer 
as can be seen from its growing density at the interface.

The relative surface excess of calcium, $\Gamma^{w}_{Ca}$, also reflects
the higher affinity of the negatively charged PC:PS (5:1) membrane compared to a neutral POPC bilayer; 
its value for the latter is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
while for the PC:PS mixture it is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
at comparable concentrations $C _{Ca} = 260\pm 30 \mathrm{mM} $.

\todo{SAMULI: This paragraph in unclear. Should be clarified and the changes
  of headgroup orientations in PC and PS clearly discussed.
}
%%% discuss the table with Gammas and follow up with molecular details, where do cations bind? Support with the spatial density figure
The density profiles of ions suggest that
the dominant contribution to the binding of \ce{Ca^{2+}} to phospholipid bilayers
comes from the interactions with the phosphate groups in both POPC and POPS. 
%The population analysis also suggests 
%that the calcium cations prefer to reside in the phosphate region of the membrane. 
%Such a finding is further validated with a more detailed analysis using Markov state modeling (MSM) \citep{Pande_MSM_paper} \todoi{Add papers reviewing MSM, e.g. recent Pande's paper.}. 
%The set of states of a calcium cation 
%encompassed all possible combinations of up to three surrounding lipids. 
%In the case of PC, we used only the phosphate moitety, which forms the dominant contribution for calcium binding.
%For PS we also distinguished configurations in which calcium interacts with the carboxylate moiety. 
%All possible combinations of such states were used to build a Markov model at a lag time $25\,\mathrm{ns}$ using pyEMMA code by \citet{pyemma} \todoi{add citation for pyemma}. 
%The resulting MSM was validated using Chapman-Kolmogorov test \citep{FrankNoe_papers_MSM} \todoi{Add citation for Noe's papers on MSM, especially Chapman Kolmogorov test.}. 
%The stationary distribution of the states reveals a strong preference of the calcium cations to reside in the phosphate region of the mixed PC-PS membrane. 
%When interacting with PS, configurations with the phosphate moiety from either PC or PS dominate the population.
%Moreover, states containing interactions with the carboxylate group in PS 
%bear higher probability when interacting also with the phosphate group of the same lipid or other lipids. 
This is also reflected in the shift of the mean orientation 
of the \ce{COO^-} group from $62^\circ$ to $73^\circ$ (420~mM \ce{CaCl2})
which was measured as the connector of the carbon atoms 
forming the bond between the group and the $\beta$-carbon of the phospholipid. 
The interactions of the carboxylate group in PS with calcium and other phosphate groups
shed light into the qualitatively different response of the head group order parameters $\alpha$ and $\beta$ in PS compared to PC 
(see Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_CaCl_PCPS}). 
We find that the complex response of the head group order parameters of POPS 
is affected by the conformational changes of the carboxylate group,
which is attracted more towards the phosphate region, 
where the calcium cations dominantly bind. 
In line with the experimental work by \citet{browning80},
this is also very likely the reason, 
why the magnitude of the P-N vector change in POPS is diminished compared to POPC, 
which is not restrained by an additional cation binding group like \ce{COO^-} in POPS.

@@

\begin{table}[tb!] 
\centering
  \caption{Bulk concentrations, $C _{Ca}$, and molar fractions, $C' _{Ca}$, of Ca$^{2+}$;
           relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
           and percentages of the population 
           of bound Ca$^{2+}$ to various moieties 
           in a neutral membrane composed of POPC
           and in a negatively charged membrane with the composition 5\,PC:1\,PS.
           \label{tab:binding}} 
  \begin{tabular}{ l | c c } 
	                     &  5\,POPC:1\,POPS &  POPC   \\
	\hline
	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
	\hline
                             &  \multicolumn{2}{c}{ } \\
        interacting moiety   &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
	\hline
	     PC              &   59   &  100   \\
	     PO$_4$    in PC &   41   &   67   \\
	     carbonyls in PC &   <1   &   ~1   \\
	\hline
	     PS              &    8   &        \\ 
	     PO$_4$  in PS   &    2   &        \\
	     COO$^-$ in PS   &    4   &        \\
	     carbonyls in PS &   <1   &        \\
	\hline
	both PC and PS       &   33   &        \\
  \end{tabular} 
\end{table}

>>>>

It has been pointed out in previous experimental works
that the conformational flexibility of PS head groups 
is different from other lipids \cite{browning80}. 
While the static shielding tensor of PS is not substantially different from PC or PE, 
the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
It is suggested that the PS head group is more rigid compared to the other phospholipids, 
which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.

We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
While the mean angle between the P-N vector of PC head group and the membrane normal
decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
 at 620~mM buffer concentration of \ce{CaCl2},
the mean orientation of the P-N vector in the PS head group has 
decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
in the same concentration range.

Such a largely diminished flexibility 
may arise from the binding of calcium to the \ce{COO^-} group in PS. 
Whe have measured the mean orientation of the \ce{COO^-} group 
as the connector of the carbon atoms 
forming the bond between the group and the $\beta$-carbon of the phospholipid. 
The mean angle between the \ce{COO^-} vector and the membrane normal
changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
 at 620~mM buffer concentration of \ce{CaCl2}.
The conformational flexibility of the PS head group is hence restrained 
by the binding of calcium cations to the \ce{COO^-} group, 
which reorients towards the membrane interior and to the phosphate region, 
where calcium cations bind predominantly.

@@

The binding of calcium cations to the phospholipid bilayer is shown 
in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
Clearly, after accounting for the electronic polarizability, 
the population of the cations in the phosphate region is significantly diminished. 
\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
Interestingly enough,
although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
 there is no extra density at the outmost surface of the membrane. 
This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
which take place during the interactions with calcium cations
described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.

The higher affinity of calcium to PS head groups compared to PC
is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
summarized in Table~\ref{tab:binding}. 
While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.

\begin{table}[tb!] 
\centering
  \caption{Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
           relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
           and percentages of the population 
           of bound Ca$^{2+}$ to various moieties 
           in a neutral membrane composed of POPC
           and in a negatively charged membrane with the composition 5\,PC:1\,PS.
           \label{tab:binding}} 
  \begin{tabular}{ l | c c } 
	                     &  5\,POPC:1\,POPS &  POPC   \\
	\hline
	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
	\hline
                             &  \multicolumn{2}{c}{ } \\
        interacting moiety   &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
	\hline
	     PC              &   59   &  100   \\
	     PO$_4$    in PC &   41   &   67   \\
	     carbonyls in PC &   <1   &   ~1   \\
	\hline
	     PS              &    8   &        \\ 
	     PO$_4$  in PS   &    2   &        \\
	     COO$^-$ in PS   &    4   &        \\
	     carbonyls in PS &   <1   &        \\
	\hline
	both PC and PS       &   33   &        \\
  \end{tabular} 
\end{table}

@@
